---
title: "Reproducible research assignment 1"
author: "Anita"
date: "05/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)

```
###Loading data###

```{r}
act <- read.csv("activity.csv")
```
###What is the average total steps taken each day? 
```{r}
total_steps <- act %>% 
 group_by(date) %>%
 summarise(steps.day = sum(steps, na.rm = TRUE))
```
##Frequency of total daily step count##
```{r}
ggplot(total_steps, aes(steps.day)) + geom_histogram(binwidth = 2000) + xlab("Total daily step count") + ylab("Frequency")
mean <- mean(total_steps$steps.day, na.rm =T) 
med <- median(total_steps$steps.day, na.rm = T) 
```
The average total daily step count 9354 and the average median number of steps taken each day is 10395


## What is the average daily activity pattern?

```{r}
###Average daily pattern###
total_steps_int <- act %>%
 group_by(interval) %>%
 summarise(steps = mean(steps, na.rm = T))

ggplot(total_steps_int, aes(x = interval, y = steps)) + geom_line() + xlab("Five minute interval") + ylab("Average step count")
total_steps_int[which.max(total_steps_int$steps),]
```
Interval 835 recorded the greatest step count per interval of 206 steps
##Imputing Missing Values##
```{r}
missing <- act[is.na(act[1]),]

#impute missing values with interval averages across days
impute_data <- act %>%
 mutate(
  steps= case_when(is.na(steps) ~ total_steps_int$steps[match(act$interval,total_steps_int$interval)], TRUE ~ as.numeric(steps))
 )
```
There are 2304 missing values in the original data set
###Plot histogram with imputed data##
```{r}
imputed <- impute_data %>% 
 group_by(date) %>% 
 summarise(im_stepsday = sum(steps))

ggplot(imputed, aes(im_stepsday)) + geom_histogram(binwidth = 2000) + xlab("Total daily step count") + ylab("Frequency")
im.mean <- mean(imputed$im_stepsday) 
im.med <- median(imputed$im_stepsday) 
diff.mean <- im.mean-mean
diff.med <- im.med-med
```
Following imputation, the mean number of daily steps is 10,766 and the median number of steps is 10,766. Imputation of missing values using the mean number of steps per interval has increased the mean and median steps by 1412 and 371 steps, respectively. 

##Are there differences in activity patterns between weekdays and weekends?
```{r}
##activity difference between weekday and weekend
library(lubridate)
Day.week <- impute_data %>% 
 mutate(
  date = ymd(date),
        day.end = case_when(wday(date) %in% 2:6 ~ "Weekday",
                            wday(date) %in% c(1,7) ~ "Weekend")
 ) %>% select(-date) %>%
 group_by(interval, day.end) %>%
 summarise(steps = mean(steps))
###total step difference between weekdays and weekends
tapply(Day.week$steps, Day.week$day.end, sum)

#Facet grid showing activity differences between weekends and weekdays
ggplot(Day.week, aes(x = interval, y = steps)) + geom_line() + facet_grid(day.end ~ .) + ylab("Average step count") + xlab("Five-minute intervals")
```
During weekdays, more steps are taken between the 500 and 1000 intervals but walking activity is greater during weekends between intervals 1000-1750 and the average step count per interval is 7 steps greater during the weekend than during weekdays. 
 

